Determining the alternatives for scalar implicature

### 2016-02-01 Paper/Paper.tex
bp pass sunday afternoon

But what are the alternatives that should be considered in the
implicature computation more generally? Under classic accounts,
listeners consider only those words whose meanings entail the actual
message (Horn, 1972), and these alternatives enter into conventionalized
or semi-conventionalized scales (Levinson, 2000). For example, because
\emph{all} entails \emph{some}, and hence is a ``stronger'' meaning,
\emph{all} should be considered as an alternative to \emph{some} in
implicatures. Similar scales exist for non-quantifier scales, e.g.
\emph{loved} entails \emph{liked} (and hence ``I liked the movie''
implicates that I didn't love it).

@@

In our current work, we pick up where these previous studies left off,
considering the set of alternatives for implicature using the RSA model.
To gain empirical traction on this issue, however, we broaden the set of
scales we consider. Our inspiration for this move comes from work by Van
Tiel, Van Miltenburg, Zevakhina, \& Geurts (2014), who examined a
phenomenon that they dubbed ``scalar diversity,'' namely the substantial
difference in the strength of scalar implicature across a variety of
scalar pairs (e.g. \emph{liked/loved,} or \emph{palatable/delicious.}).
Making use of some of this diversity allows us to investigate the ways
that different alternative sets give rise to implicatures of different
strengths across scales.

@@

\[p_{S_1}(u \mid m) \propto p_{L_0} (m \mid u)\]

@@

\[p_{S_1}(u \mid m) \propto e^{-\alpha[-log(p_{L_0}(m \mid u)) - C(u)]}\]

@@

To instantiate our signaling game with a tractable message set \(M\), in
our studies we adopt the world of restaurant reviews as our
communication game. We assume that speakers and listeners are trying to
communicate the number of stars in an online restaurant review (where
\(m \in \{1, 2, 3, 4, 5\}\)). We then use experiments to measure three
components of the model. First, to measure literal semantics
\({p_{L_0} (m \mid u)}\) (we ask experiment participants to judge
whether a message is compatible with a particular meaning (Experiment
1). Second, to generate a set of plausible alternative messages in
\(U\), we elicit alternatives directly (Experiment 2). Lastly, to obtain
human \(L_1\) pragmatic judgments, we ask participants to interpret a
speaker's utterances (Experiment 3).

@@

Experiment 1 was conducted to approximate literal listener semantic
distributions \(p_{L_0}(m \mid u)\) for five pairs of scalar items taken
from Tiel (2014). We include three conditions in Experiment 1,
corresponding to the sets of alternatives within a scale that
participants were presented with: two alternatives (``entailment''),
four alternatives, and five alternatives. The two alternatives condition
makes a test of the hypothesis that the two members of the classic Horn
(entailment) scale (Horn, 1972) are the only alternatives necessary to
predict the strength of listeners' pragmatic inference. The four and
five alternatives conditions then add successively more alternatives to
test whether including a larger number of alternatives will increase
model
fit.\footnote{Note that alternatives in the four and five alternatives conditions were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; all literal listener experiments are grouped together for simplicity in reporting.}
A secondary goal of Experiment 1 is to test whether the set of
alternatives queried during literal semantic elicitation impacts
compatibility judgments. (If it does we should see differences in
compatibility judgments for shared items between experiments.)

@@

Conditions were run sequentially. In each condition we recruited 30
participants from Amazon Mechanical Turk (AMT). In the two alternative
condition, 16 participants were excluded for either failing to pass two
training trails or were not native English speakers, leaving a total
sample of 14
participants.\footnote{The majority of respondent data excluded from the two alternative condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.}
In the four alternative condition, 7 participants were excluded for
either failing to pass two training trials or were not native English
speakers, leaving a total sample of 23 participants. In the five
alternative condition, 3 participants were excluded for either failing
to pass two training trials or were not native English speakers, leaving
a total sample of 27.

@@

The two alternatives condition included only the scalar pairs from Tiel
(2014). The four alternatives condition included the two scalar items
plus the top two alternatives generated for each scalar family by
participants in Experiment 2. The five alternatives condition included
the four previous items plus one more neutral item chosen from those
alternatives generated in Experiment 2. See Figure
\ref{fig:allScalesTable} for the complete list of alternatives used in
each condition.

@@

Figure \ref{fig:exp2Plots} plots literal listener \(p_{L0}(m|u)\) scalar
item distributions for the three conditions. Each row shows a unique
scalar family with items ordered horizontally by valence. Several trends
are visible. First, in each scale, the alternatives spanned the star
scale, such that there were alternatives that were highly compatible
with both the lowest and highest numbers of stars. Second, there was
clear variability between scalar families. For example, compatibility
judgments for the top items in the \emph{memorable / unforgettable}
scale were more similar than those for \emph{good / excellent} or
\emph{liked / loved}. Finally, there was substantial consistency in
ratings for items that were repeated across experiments, suggesting that
this paradigm elicited stable judgments from participants.

To test our secondary hypothesis, that the different sets of scalar
items might result in differences in compatibility judgments between
experiments, we ran a mixed effects model. We regressed compatibility
judgments on scale, number of stars and condition, with subject and word
level random effects, which was the maximal structure that converged.
Results indicate no significant differences between Experiment 1a and
1c, \(b =\)-0.05, \(Z =\) -0.53, \(p =\) 0.59 or between 1b and 1c,
\(b =\) -0.04, \(Z =\) -0.52 \(p =\) 0.6 and the addition of condition
as a predictor did not significantly improve model fit when compared to
a model without the experiment variable using ANOVA \(\chi^2(\) 2
\() =\) 0.43, \(p =\) 0.81.

@@

To elicit empirical alternatives for the scales we used in Experiment 1,
we adopted a modified cloze task inspired by Experiment 3 of Tiel
(2014).

@@

\caption[Combined counts for participant-generated alternatives for the 'liked  loved' scale in Experiment 2]{Combined counts for participant-generated alternatives for the 'liked  loved' scale in Experiment 2.}\label{fig:exp2_altsPlot_likedLoved}
\end{figure}
\end{CodeChunk}

@@

\caption[Pragmatic listener judgements for scalar items]{Pragmatic listener judgements for scalar items. Proportion of participants generating a star rating is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Line type denotes condition, and colors indicate the particular scalar items. Each panel shows one scalar pair, with only entailment items (two alternatives condition) shown here for simplicity.}\label{fig:exp2Plots}
\end{figure*}
\end{CodeChunk}

Figure \ref{fig:exp2_altsPlot_likedLoved} shows an example alternative
set for the scalar items \emph{liked} and \emph{loved} (combined).
Alternative distributions for the other scalar pairs (e.g..
\emph{good/excellent}, \emph{memorable/unforgettable}) were similarly
long-tailed.

@@

Experiment 3 was conducted to measure pragmatic
judgments---\(p_{L_1}(m \mid u)\) in Equation \((1)\). As in Experiment
1, we include several conditions to test inferences in the presence of
different alternative sets. In the two alternatives condition,
participants made judgments for items included in the entailment scales.
In the four alternatives condition, participants made judgments for the
entailment items and also the top two alternatives elicited for each
scale in Experiment 2. Including two conditions with differing
alternatives allowed us to rule out the potential effects of having a
larger set of alternatives during the pragmatic judgment elicitation and
also provided two sets of human judgments to compare with model
predictions.\footnote{Note that alternatives in the four alternatives condition were chosen on the basis of Experiment 2, which was run chronologically after the two alternatives condition; both pragmatic listener experiments are grouped together for simplicity in reporting.}

@@

We recruited 100 participants from AMT, 50 for each condition. Data for
9 participants was excluded from the two alternatives condition after
participants either failed to pass two training trials or were
non-native English speakers, leaving a total sample of 41 participants.
In the four alternative condition, data from 7 participants was excluded
after participants either failed to pass two training trials or were not
native English speakers, leaving 43 participants.

@@

To test our secondary hypothesis, that the different sets of scalar
items in the two conditions might result in differences in pragmatic
judgments for the same items, we fit a mixed effects model. We regressed
pragmatic judgments on scale and condition with subject- and word-level
random effects, which was the maximal structure that converged. There
were no significant differences between conditions, \(b =\) 0.05, \(t(\)
150 \() =\) 1.04, \(p =\) 0.3 and the addition of the condition
predictor did not significantly improve model fit when compared to a
model without that variable \(\chi^2(\) 1 \() =\) 1.13, \(p =\) 0.29.

@@

Using literal listener data from Experiment 1, we conducted a set of
simulations with the RSA model. Each simulation kept the model constant,
fitting the choice parameter \(\alpha\) as a free parameter, but used a
set of alternatives to specify the scale over which predictions were
computed. We considered four different alternative sets, with empirical
measurements corresponding to those shown in Figure
\ref{fig:allScalesTable}: 1) the two alternatives in the classic
entailment scales, 2) those two alternatives with the addition of a
generic negative alternative, 3) the expanded set of four alternatives
tested in Experiment 1, and 4) the expanded set of five alternatives
tested in Experiment 1.

Model fit with human judgments was significantly improved by the
inclusion of alternatives beyond the entailment items (Table 1). The two
alternatives model contained only entailment items, which, under classic
accounts, should be sufficient to generate implicature, but fit to data
was quite poor with these items. The addition of a generic negative
element produced some gains in performance, but much higher performance
was found when we included four and five alternatives, with the
alternatives derived empirically for the specific scale we used. An
example fit for the five-alternative model is shown in Figure
\ref{fig:fiveAltsScatter}.

@@

\caption[Judgments from Experiment 3, four alternatives condition, are plotted against model predictions]{Judgments from Experiment 3, four alternatives condition, are plotted against model predictions. Colors show the star rating for individual judgments.}\label{fig:fiveAltsScatter}
\end{figure*}
\end{CodeChunk}

@@

While improvements in model fit were substantial as we moved from two to
four alternatives, we saw only a minor increase in fit from the move to
five alternatives. One possible explanation is that alternatives are
differentially salient in context, and moving to larger sets would
consider weighting the alternatives differentially (as Franke, 2014
did). Preliminary simulations using weightings derived from Experiment 2
provide some support for this idea but would require further empirical
work for confirmation.

>>>>

But what are the alternatives that should be considered in the
implicature computation more generally? Under classic accounts,
listeners consider only those words whose meanings entail the actual
message (Horn, 1972), and these alternatives enter into conventionalized
or semi-conventionalized scales (Levinson, 2000). For example, because
\emph{all} entails \emph{some}, and hence is a ``stronger'' meaning,
\emph{all} should be considered as an alternative to \emph{some} in
implicatures. Similar scales exist for non-quantifier scales, e.g.
\emph{loved} entails \emph{liked} (and hence ``I liked the movie''
implicates I didn't love it).

@@

In our current work, we pick up where these previous studies left off,
considering the set of alternatives for implicature using the RSA model.
To gain empirical traction on this issue, however, we broaden the set of
scales we consider. Our inspiration for this move comes from work by Van
Tiel, Van Miltenburg, Zevakhina, \& Geurts (2014), who examined a
phenomenon that they dubbed ``scalar diversity,'' namely the substantial
difference in the strength of scalar implicature across a variety of
scalar pairs (e.g. \emph{liked}/\emph{loved,} or
\emph{palatable}/\emph{delicious.}). Making use of some of this
diversity allows us to investigate the ways that different alternative
sets give rise to implicatures of different strengths across scales.

@@

\[p_{S_1}(u \mid m) \propto p_{L_0} (m \mid u) \tag{2}\]

@@

\[p_{S_1}(u \mid m) \propto e^{-\alpha[-log(p_{L_0}(m \mid u)) - C(u)]} \tag{3}\]

@@

To instantiate our signaling game with a tractable message set \(M\), in
our studies we adopt the world of restaurant reviews as our
communication game. We assume that speakers and listeners are trying to
communicate the number of stars in an online restaurant review (where
\(m \in \{1, 2, 3, 4, 5\}\)). We then use experiments to measure three
components of the model. First, to measure literal semantics
\({p_{L_0} (m \mid u)}\) we ask experiment participants to judge whether
a message is compatible with a particular meaning (Experiment 1).
Second, to generate a set of plausible alternative messages in \(U\), we
elicit alternatives directly (Experiment 2). Lastly, to obtain human
\(L_1\) pragmatic judgments, we ask participants to interpret a
speaker's utterances (Experiment 3).

@@

Experiment 1 was conducted to approximate literal listener semantic
distributions \(p_{L_0}(m \mid u)\) for five pairs of scalar items taken
from Tiel (2014). We include three conditions in Experiment 1,
corresponding to the sets of alternatives within a scale that
participants were presented with: two-alternatives (``entailment''),
four-alternatives, and five-alternatives. The two-alternatives condition
makes a test of the hypothesis that the two members of the classic Horn
(entailment) scale (Horn, 1972) are the only alternatives necessary to
predict the strength of listeners' pragmatic inference. The four and
five-alternatives conditions then add successively more alternatives to
test whether including a larger number of alternatives will increase
model
fit.\footnote{Note that alternatives in the four and five-alternatives conditions were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; all literal listener experiments are grouped together for simplicity in reporting.}
A secondary goal of Experiment 1 is to test whether the set of
alternatives queried during literal semantic elicitation impacts literal
semantic judgments. If it does we should see differences in these
judgments for shared items between experiments.

@@

Conditions were run sequentially. In each condition we recruited 30
participants from Amazon Mechanical Turk (AMT). In the two-alternatives
condition (Experiment 1a), 16 participants were excluded for either
failing to pass two training trails or were not native English speakers,
leaving a total sample of 14
participants.\footnote{The majority of respondent data excluded from the two-alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.}
In the four alternative condition (Experiment 1b), 7 participants were
excluded for either failing to pass two training trials or were not
native English speakers, leaving a total sample of 23 participants. In
the five alternative condition (Experiment 1c), 3 participants were
excluded for either failing to pass two training trials or were not
native English speakers, leaving a total sample of 27.

@@

The two-alternatives condition included only the scalar pairs from Tiel
(2014). The four-alternatives condition included the two scalar items
plus the top two-alternatives generated for each scalar family by
participants in Experiment 2. The five-alternatives condition included
the four previous items plus one more neutral item chosen from those
alternatives generated in Experiment 2. See Figure
\ref{fig:allScalesTable} for the complete list of alternatives used in
each model.

@@

Figure \ref{fig:exp2Plots} plots literal listener \(p_{L0}(m|u)\) scalar
item distributions for the three conditions. Each row shows a unique
scalar family with items ordered horizontally by valence. Several trends
are visible. First, in each scale, the alternatives spanned the star
scale, such that there were scalar items that were highly compatible
with both the lowest and highest numbers of stars. Second, participant
compatibility judgments match intuitive literal semantics. That is, both
weaker (e.g. \emph{some} or \emph{good}) and stronger (e.g. \emph{all}
or \emph{excellent}) entailment items were seen as compatible with
five-star ratings. This means participants' compatibility judgements
reflect literal semantic intuitions, not enriched pragmatic judgments.
We see clear variability between scalar families (literal semantic
distributions for \emph{memorable} are considerably different from
\emph{palatable}), but also substantial consistency across conditions
(compatibility judgments for individual scalar items are consistent in
Experiments 1a,b,c).

To examine this last issue of consistency across experiments (our
secondary hypothesis) we ran a mixed effects model regressing
compatibility judgments on scale, number of stars and experimental
condition, with subject and word level random effects, which was the
maximal structure that converged. Results indicate no significant
differences between Experiment 1a and 1c, \(b =\)-0.05, \(Z =\) -0.53,
\(p =\) 0.59 or between 1b and 1c, \(b =\) -0.04, \(Z =\) -0.52 \(p =\)
0.6 and the addition of condition as a predictor did not significantly
improve model fit when compared to a model without the experiment
condition variable using ANOVA \(\chi^2(\) 2 \() =\) 0.43, \(p =\) 0.81.
These findings suggest that literal semantic intuitions are not effected
by the set of alternatives queried in our paradigm. This finding is
important because literal semantic data generated in these three
conditions are used to simulate the effects of different alternative
sets on implicature generation in our model.

@@

To examine the effect different alternative sets have on implicature, we
needed a way of selecting plausible alternatives. We addressed this
issue by adopting a modified cloze task to elicit empirical alternatives
for the entailment items we used in Experiment 1a and 3a. This design
was inspired by Experiment 3 of Tiel (2014).

@@

\caption[Combined counts for participant-generated alternatives for the `liked  loved' scale in Experiment 2]{Combined counts for participant-generated alternatives for the `liked  loved' scale in Experiment 2.}\label{fig:exp2_altsPlot_likedLoved}
\end{figure}
\end{CodeChunk}

@@

\caption[Pragmatic listener judgements for scalar items]{Pragmatic listener judgements for scalar items. Proportion of participants generating a star rating is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Line type denotes condition, and colors indicate the particular scalar items. Each panel shows one scalar pair, with only entailment items (two-alternatives condition) shown here for simplicity.}\label{fig:exp2Plots}
\end{figure*}
\end{CodeChunk}

Figure \ref{fig:exp2_altsPlot_likedLoved} shows an example alternative
set for the scalar items \emph{liked} and \emph{loved} (combined). We
expanded our alternative set from only entailment items (Experiments 1a,
3a) to four alternatives (Experiments 1b, 3b) by including the top two
alternatives elicited for each scale in Experiment 2. For example, the
top two alternatives for the \emph{liked}/\emph{loved} scale were
\emph{hated} and \emph{disliked} and were included in four-alternatives
model runs. Similarly, neutral items (e.g. \emph{okay} in the
\emph{good}/\emph{excellent} scalar family) were chosen from among the
set of alternatives elicited in Experiment 2. Alternative set
distributions for the other scalar families (e.g.
\emph{good}/\emph{excellent}, \emph{memorable}/\emph{unforgettable})
were similarly long-tailed. See Figure \ref{fig:allScalesTable} for the
complete list of alternatives we used from Experiment 2.

@@

Experiment 3 was conducted to measure pragmatic
judgments---\(p_{L_1}(m \mid u)\) in Equation \((1)\). As in Experiment
1, we include several conditions to test inferences in the presence of
different alternative sets. In the two-alternatives condition
(Experiment 3a), participants made judgments for items included in the
entailment scales. In the four-alternatives condition (Experiment 3b),
participants made judgments for the entailment items and also the top
two-alternatives elicited for each scale in Experiment 2. Including two
conditions with differing alternatives allowed us to rule out the
potential effects of having a larger set of alternatives during the
pragmatic judgment elicitation and also provided two sets of human
judgments to compare with model
predictions.\footnote{Note that alternatives in the four-alternatives condition were chosen on the basis of Experiment 2, which was run chronologically after the two-alternatives condition; both pragmatic listener experiments are grouped together for simplicity in reporting.}

@@

We recruited 100 participants from AMT, 50 for each condition. Data for
9 participants was excluded from the two alternatives condition after
participants either failed to pass two training trials or were
non-native English speakers, leaving a total sample of 41 participants.
In the four alternatives condition, data from 7 participants was
excluded after participants either failed to pass two training trials or
were not native English speakers, leaving 43 participants.

@@

To rule out the potential effects of having a larger set of alternatives
during the pragmatic judgment elicitation, we fit a mixed effects model.
We regressed pragmatic judgments on scale and experimental condition
with subject- and word-level random effects, which was the maximal
structure that converged. There were no significant differences between
conditions 3a and 3b, \(b =\) 0.05, \(t(\) 150 \() =\) 1.04, \(p =\) 0.3
and the addition of the condition predictor did not significantly
improve model fit when compared to a model without that variable
\(\chi^2(\) 1 \() =\) 1.13, \(p =\) 0.29.

@@

Using literal listener data from Experiment 1, we conducted a set of
simulations with the RSA model. Each simulation kept the model constant,
fitting the choice parameter \(\alpha\) as a free parameter, but used a
set of alternatives to specify the scale over which predictions were
computed. We considered four different alternative sets, with empirical
measurements corresponding to those shown in Figure
\ref{fig:allScalesTable}: 1) the two-alternatives in the classic
entailment scales, 2) those two alternatives with the addition of a
generic negative alternative
\footnote{literal semantics for the generic negative alternative were not generated empirically, but serve as a baseline "none" style semantics in which the scalar item is only compatible with 1 star},
3) the expanded set of four alternatives tested in Experiment 1b, and 4)
the expanded set of five alternatives tested in Experiment 1c.

Model fit with human judgments was significantly improved by the
inclusion of alternatives beyond the entailment items (Table 1). The
two-alternatives model contained only entailment items, which, under
classic accounts, should be sufficient to generate implicature, but fit
to data was quite poor with these items. The addition of a generic
negative element produced some gains in performance, but much higher
performance was found when we included four and five alternatives, with
the alternatives derived empirically for the specific scale we used. An
example fit for the five-alternative model is shown in Figure
\ref{fig:fiveAltsScatter}.

@@

\caption[Judgments from Experiment 3, four-alternatives condition, are plotted against model predictions]{Judgments from Experiment 3, four-alternatives condition, are plotted against model predictions. Colors show the star rating for individual judgments.}\label{fig:fiveAltsScatter}
\end{figure*}
\end{CodeChunk}

@@

While improvements in model fit were substantial as we moved from two to
four alternatives, we saw only a minor increase in fit from the move to
five alternatives. One possible explanation is that alternatives are
differentially salient in context, and in moving to larger sets we
should consider weighting the alternatives differentially (as Franke,
2014 did). Preliminary simulations using weightings derived from
Experiment 2 provide some support for this idea but would require
further empirical work for confirmation.
